Can We Crunch the Social Graph on the Cheap?

### 2011-01-25 doc/hotpar2011/softxmt-hotpar2011.tex
hardware related work

\section{Related work}
\label{sec:related}

@@

Multithreading for latency tolerance has been explored in a number of directions.

Hardware approaches: The Tera MTA and Cray XMT, the MIT Alewife
project, ??? use hardware context switches to tolerate memory latency
for large systems. More recently, Simultaneous Multi-Threading,
Niagara, Cyclops, and even GPUs switch between multiple contexts to
tolerate memory latency, for memory that is hundreds of cycles of
away, rather than the thousands we must tolerate.

software:

Mowry's SCM

Cullers' TAM

qthreads

cilk? tbb?

why not wait for more cores?

>>>>

\section{Related work}
\label{sec:related}
The concept of using independent parallel contexts to tolerate latency
is not novel.  The inspiration for our project is of course the Cray
XMT \cite{feo-xmt}, which implemented a hardware version of our
programming model (in what would today be considered exceptionally
slow silicon.)  The XMT can be seen as simultaneous multithreading
\cite{tullsen-smt} writ large; the technique and motivation is
essentially the same, but the XMT allows hundreds of contexts (modern
Intel SMT implementations are generally limited to two.) The Alewife
system \cite{agarwal-alewife} also switches contexts (in hardware) to
hide latency on remote-node accesses.

Modern programmable GPUs can also be interpreted in using an XMT-like
model under the hood;
% Do I need to include a citation here?
while they generate large numbers of available execution contexts via
a (semi-)explicit SIMD model, those SIMD instructions keep functional
units satisfied by overlapping execution with the loads required for
other contexts.

Similar techniques have also been used in software.  Mowry proposed a system \cite{mowry-scm} of \emph{software-controlled
  multithreading}.  As opposed to our system of explicitly marked
remote references, SCM assumes a high hit rate and traps to an
interrupt routine which performed a context switch upon a cache miss.  The 2000-era processors had
short enough latencies that using more than two contexts generally
produced slowdown.
 b. TAM
 c. qthreads
Of course, our system is proceeded by the hardware
\emph{software controlled multithreading},

@@
